---
title: "Lab 6 Binary Outcome"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Importing your data 

Let's first load the data `WAGE2.dtt`

```{r mlr}

# load your data and call it data 
library(haven)
data = read_dta('WAGE2.dta')

# Descriptive statistics and the overview of the data 

library(psych)
#describe(data)


# Let's plot the wage distribution (`wage`)

```

### Let's create a multiple regression model 

- DV: wage_binary (will be modified to 1=high income, 0=low income)
- IV: educ sibs meduc feduc 

```{r mlr2}
# Let's get an average wage first 
mean(data$wage)
```

### Assumption Checking 
# Assuming model_logit is already fitted with glm(..., family = binomial)
# and model_data is your cleaned dataset (no missing values)
```{r}
# 1. Check Binary Outcome
# wage > average (1=high income) otherwise low income 
data$wage_binary <- ### YOUR CODE HERE ###
  
table(model_data$wage_binary)  # Should have two levels: 0 and 1

# 2. Multicollinearity - VIF should be < 5 (or < 10 as a relaxed rule)
library(car)
vif(model_logit)

# 3. Influential observations - Cook's distance and leverage
influence_measures <- influence.measures(model_logit)

# Plot Cook’s Distance
plot(model_logit, which = 4, cook.levels = 0.5)  # Default cook.levels = 0.5

# Identify all the influential points
cooksD <- cooks.distance(model_logit)
influential <- which(cooksD > (4 / nrow(model_data)))
influential

```

## model building (1) - original data
```{r}
model_logit <- glm(wage_binary ~ educ + sibs + meduc + feduc, 
                   data = data, 
                   family = binomial(link = "logit")) 

summary(model_logit)

```

## model building (2) - remove the influential points
```{r}
model_logit_no_outliers <- glm(wage_binary ~ educ + sibs + meduc + feduc, 
                               data = model_data[-influential, ], 
                               family = binomial)
summary(model_logit_no_outliers)

```

## evaluate the coefficients 
Be aware, that the output of the logistic model is on link-scale (logit). Thus, the numerical output of the model corresponds to the log-odds. For convenience let us write down the logistic regression model with the calculated intercept and coefficient:

```{r test1}
summary(model_logit)$coefficients

```

```{r test2}
exp(coefficients(model_logit))

```

#### confidence interval (coefficients )

```{r robustse}
confint.default(model_logit)
```

## Goodness-of-Fit    
```{r q4}
deviance(model_logit)
null.deviance <- with(model_logit, null.deviance)
residual.deviance <- with(model_logit, deviance)
paste("Null deviance:", null.deviance)
paste("Residual deviance:", residual.deviance)

#Pseudo R-squared score 
#install.packages('pscl')
library(pscl)
pR2(model_logit)
```

## Model Evaluation: Confusion Matrix, Precision, Recall

```{r evaluation}
# Remove rows with missing values used in the model
model_data <- na.omit(data[, c("wage_binary", "educ", "sibs", "meduc", "feduc")])

# Fit model again with cleaned data
model_logit <- glm(wage_binary ~ educ + sibs + meduc + feduc, 
                   data = model_data, 
                   family = binomial(link = "logit"))

# Generate predicted probabilities
pred_probs <- predict(model_logit, type = "response")

# Apply threshold of 0.5
pred_class <- ifelse(pred_probs >= 0.5, 1, 0)

# Convert to factor for caret
pred_class <- factor(pred_class, levels = c(0,1))
actual_class <- factor(model_data$wage_binary, levels = c(0,1))

# Load caret and compute confusion matrix
#install.packages('caret')
library(caret)
conf_mat <- confusionMatrix(pred_class, actual_class, positive = "1")
conf_mat

# Precision and recall
conf_mat$byClass["Precision"]
conf_mat$byClass["Recall"]
```

## Plotting your results 
- This shows how each predictor affects the predicted probability, holding other variables constant.
```{r}
#install.packages('effects')
# install.packages("effects")
library(effects)

plot(allEffects(model_logit))  # This gives partial effect plots for each predictor
```

#########################
### Classrom Exercise ###
#########################
In this section, you will:
> 1. Add an interaction term to your logistic regression model
> 2. Interpret its effect on the probability of being in the high-income group

```{r}
# Step 1: Fit a Logistic Regression Model with Interaction between educ and meduc







# Is the interaction effect statistically significant? 
# What does the sign of the coefficient suggest? 

``` 
